Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2018-10-04 paper/SI.tex
Minor changes to main and SI

\title{Macroscopic Salt Rejection through Electrostatically Gated Monolayer Porous Graphene}

@@

\section{Ion diffusion measurement and calculation}
\label{sec:measure}

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    {[Fe(CN)$_{6}$]$^{3-}$}   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical diffusion through PCTE can be
    calculated using Eq. 1 in the main text and the diffusivities of
    the respective anion and cation.}
  \label{tab:diff}
\end{table}

@@

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 0.0151 $\pm$ 0.00265  & 106   & 1.34 $\pm$ 0.235\\
    NaCl        & 0.0099 $\pm$ 0.000173 & 91    & 1.02 $\pm$ 0.0179\\
    LiCl        & 0.00603 $\pm$ 0.000802        & 83    & 0.679 $\pm$ 0.0903\\
    CaCl$_{2}$  & 0.0125 $\pm$ 0.001    & 149   & 0.789 $\pm$ 0.0631\\
    MgSO$_{4}$  & 0.0088 $\pm$ 0.0012   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}$] & 0.024 $\pm$ 0.00275   & 395   & 0.575 $\pm$ 0.0633\\
    K$_{2}$SO4$_{4}$    & 0.0198 $\pm$ 0.00199  & 258   & 0.722 $\pm$ 0.0726\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

\section{Control experiment with copper tape}
\label{sec:copper}
We make sure that the applied voltage is effectively applied via the
graphene membrane and not through leakage directly coupling the
coppertape to the ionic solution. For this, we use a device as in
\Fig 1c \todo[inline]{which figure?}, omitting the graphene and
measuring the current through the membrane with the Autolab
electrochemical workstation using the membrane as working electrode
and a platinum wire as counter/reference electrode 3.5 cm apart in 0.1
mM KCl. When no graphene is inserted in the device, a current in the
baseline range ($\sim$ 0.1 nA) is observed, indicating no current
passing from the membrane. However, if graphene is inserted, a current
in the range of $\sim$ 10 - 100 nA is measured. If the copper tape is
in contact with the solution when omitting the Kapton tape, the
current increases to $\sim$ 1 - 10 $\mathrm{\mu}$A, independent of the
exposed area of the copper tape to the solution.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lcc}
    \hline
    Solution & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm) \\
    \hline
    0.1 mM NaCl &0.018  &30.4\\
    0.1 mM LiCl &0.0158 &30.4\\
    0.1 mM CaCl$_{2}$&  0.02187 &17.2\\
    0.1 mM MgSO$_{4}$   &0.0327 &15.2\\
    0.1 mM K$_{2}$SO$_{4}$      &0.0416 &17.2\\
    0.1 mM K$_{3}${[Fe(CN)$_{6}$]}&       0.0595  &9.6\\
    \hline
  \end{tabular}
  \caption{Conductivities and Debye lengths for the 6 different salts
    under investigation.}
  \label{tab:conductivity}
\end{table}

>>>>

\title{Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene}

@@

\section{Supplementary Section 2.       Ion diffusion measurement and calculation}
\label{sec:measure}

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    Fe{[(CN)$_{6}$]}$^{3-}$   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical
    diffusion through PCTE can be calculated using Eq. 1 in the main
    text and the diffusivities of the respective anion and cation.}
  \label{tab:diff}
\end{table}

@@

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 0.0151 $\pm$ 0.00265  & 106   & 1.34 $\pm$ 0.235\\
    NaCl        & 0.0099 $\pm$ 0.000173 & 91    & 1.02 $\pm$ 0.0179\\
    LiCl        & 0.00603 $\pm$ 0.000802        & 83    & 0.679 $\pm$ 0.0903\\
    CaCl$_{2}$  & 0.0125 $\pm$ 0.001    & 149   & 0.789 $\pm$ 0.0631\\
    MgSO$_{4}$  & 0.0088 $\pm$ 0.0012   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 0.024 $\pm$ 0.00275   & 395   & 0.575 $\pm$ 0.0633\\
    K$_{2}$SO4$_{4}$    & 0.0198 $\pm$ 0.00199  & 258   & 0.722 $\pm$ 0.0726\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

\section{Control experiment with copper tape}
\label{sec:copper}
We make sure that the applied voltage is effectively applied via the
graphene membrane and not through leakage directly coupling the
coppertape to the ionic solution. For this, we use a device as in \Fig
1a, omitting the graphene and measuring the current through the
membrane with the Autolab electrochemical workstation using the
membrane as working electrode and a platinum wire as counter/reference
electrode 3.5 cm apart in 0.1 mM KCl. When no graphene is inserted in
the device, a current in the baseline range ($\sim$ 0.1 nA) is
observed, indicating no current passing from the membrane. However, if
graphene is inserted, a current in the range of $\sim$ 10 - 100 nA is
measured. If the copper tape is in contact with the solution when
omitting the Kapton tape, the current increases to $\sim$ 1 - 10
$\mathrm{\mu}$A, independent of the exposed area of the copper tape to
the solution.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lcc}
    \hline
    Solution & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm) \\
    \hline
    0.1 mM NaCl &0.018  &30.4\\
    0.1 mM LiCl &0.0158 &30.4\\
    0.1 mM CaCl$_{2}$&  0.02187 &17.2\\
    0.1 mM MgSO$_{4}$   &0.0327 &15.2\\
    0.1 mM K$_{2}$SO$_{4}$      &0.0416 &17.2\\
    0.1 mM K$_{3}$[Fe(CN)$_{6}$]&       0.0595  &12.4\\
    \hline
  \end{tabular}
  \caption{conductivity}
  \label{tab:conductivity}
\end{table}
